Leveraging Altruistic Peers to Reduce the Bandwidth Costs of Free Content Downloads

### 2008-05-02 docs/abstract/abstract.tex
Final abstract for submission.

\title{Using Altruistic Peers to Help With Free Content Downloads}
\author{\IEEEauthorblockN{Cameron Dale}
\IEEEauthorblockA{School of Computing Science\\
Simon Fraser University\\
Burnaby, British Columbia, Canada\\
Email: camerond@cs.sfu.ca}
\and
\IEEEauthorblockN{Jiangchuan Liu}
\IEEEauthorblockA{School of Computing Science\\
Simon Fraser University\\
Burnaby, British Columbia, Canada\\
Email: jcliu@cs.sfu.ca}}

@@

% \begin{abstract}
% A large amount of free content is available over the Internet from
% many different distributors. Most of this content uses the traditional
% client-server model to handle requests from users. However, there is an
% opportunity to use peer-to-peer techniques to reduce the cost of
% much of this distribution, especially due to the altruistic nature
% of many of these users. We present a general technique for
% satisfying the needs of this P2P distribution, which is applicable
% to many of these distributors systems. Our method makes use of a DHT
% for storing the location of peers, using the cryptographic hash of
% the file as a key. We then go further and implement a solution for
% the distribution of Debian software packages.
% 
% We have designed a generally applicable peer-to-peer content
% distribution model to be used by many of the free content
% distributors operating today. It makes use of already existing
% features of most package management systems, to create a
% peer-to-peer distribution that should substantially reduce the costs
% of hosting the content.
% 
% We have also implemented our design in software to be used in
% conjuction with Debian-based distribution of Linux software
% packages. It will soon be in use by many users of the Debian
% project's distribution, and so serves as an example of the
% opportunities that exist.
% \end{abstract}

@@

There are a large number of free content distributors using custom
package distribution systems over the Internet to distribute content
to their users. These distributors have developed many different
methods for this distribution, but they almost exclusively use a
client-server model to satisfy user requests. The contents' size,
popularity and number of users results in a large number of
requests, which usually requires a network of mirrors to handle. Due
to the free nature of this content, many users are willing and able
to contribute upload bandwidth to this distribution, but currently
have no way to do this.

There are many peer-to-peer implementations available today, but
none is very well suited to this specific problem. Many distributors
make their content available in some form using BitTorrent, but it
is not ideally suited to an application such as this one. BitTorrent
is designed for large static collections of files, in which all
peers are interested in most or all of the content. But the packages
in this case are small, constantly being updated, and users are only
interested in a small subset of them. Other proposed implementations
suffer from some of the same, and other, problems, and so none has
been well-used to address this problem.

We present a new peer-to-peer distribution model to meet these
demands. It is based on the lessons learned from many previous
implementations of successful peer-to-peer protocols, and makes use
of a distributed hash table (DHT) and borrows many ideas from
BitTorrent. The model relies on the pre-existence of cryptographic
hashes of the content, which uniquely identify it for a request from
other peers. If the peer-to-peer download fails, then the original
request to the server is used as a fallback to prevent dissatisfied
users. The peer can then share this new content with others through
the P2P system. This system would efficiently satisfy the demands of
a large number of users, and should significantly reduce the
currently substantial bandwidth costs of hosting this content.

We also present a sample implementation based on the Debian package
distribution system. This implementation is simple, makes use of
mostly well-known technologies, and so serves as an example for
other free content distributors of the opportunity that can be met
with such a system.

@@

There are a large number of groups using the Internet to distribute
their free content. This content is made available from a free
download site, which usually requires a network of mirrors to
support the large number of requests. This content almost always
supports some kind of file verification, using cryptographic hashes,
to verify the completed downloads accuracy or authenticity.

Most Linux distributions use a software package management system
that fetches packages to be installed from a network of mirrors.
Debian-based distributions uses the \texttt{apt} program, which
downloads Debian packages from one of many mirrors. RPM-based
distributions use \texttt{yum}, and Gentoo uses \texttt{portage},
which both operate in a similar way. Other free software
distributors also use a similar system: CPAN distributes files
containing software packages for the PERL programming language,
using SOAP RPC requests to find and download files; Cygwin provides
many of the standard Unix/Linux tools in a Windows environment,
using a package management tool that requests packages from
websites; there are two software distribution systems for Mac OSX,
fink and MacPorts, that also retrieve packages in this way. Also,
some systems use direct web downloads, but with a hash verification
file also available for download next to the desired file. These
hash files usually have the same file name, but with an added
extension identifying the hash used (e.g. \texttt{.md5} for the MD5
hash). Finally, there are other programs that make use of
cryptographic hashing to identify files: e.g. Git is a version
control system in which all files and changes are identified by
their SHA1 hash.

These systems all share some commonalities: the content is avaliable
for anyone to freely download, the content is divided into distinct
units (packages), users are typically not interested in downloading
all of the content available, and hashes of the content are
available before the download is attempted. We also expect that
there are a number of users of these systems that are motivated by
altruism to want to help out with this distribution. This is common
in these systems due to the free nature of the content being
delivered.

@@

This situation presents a clear opportunity to use some form of
peer-to-peer file-sharing protocol. The sparse interest in a large
number of packages undergoing constant updating is well suited to
the functionality provided by a Distributed Hash Table (DHT). DHTs
require unique keys to store and retrieve strings of data, which the
cryptographic hashes used by these package management systems are
perfect for. The stored and retrieved strings can then be pointers
to the peers that have the content package that hashes to that key.
A downloading peer can lookup the package hash in the DHT and, if it
is found, download the file from those peers and verify the content.
If the package hash can not be found in the DHT, the peer will
fallback to downloading from the original content location, and once
complete will add a new entry to the DHT indicating that it now has
the content. The servers or mirrors thus act as \emph{seeds} for the
P2P system without any modification to them.

This desired P2P functionality could be integrated into the existing
package management software in 2 ways. The functionality can be
directly integrated into the software, though this could be
difficult as the DHT should be running at all times, whereas the
tools typically only run until the download request is complete.
Alternatively, since many of the package management software
implementations use HTTP requests to download the files, this makes
it possible to implement the P2P aspect as a standard HTTP caching
proxy, which will get uncached requests first from the P2P system,
and then fallback to the normal HTTP request from a server. For
methods that don't use HTTP requests, other types of proxies may
also be possible.

Downloading a large file efficiently from a number of peers is where
BitTorrent shines as a peer-to-peer application. Its method of
breaking up larger files into pieces, each with its own hash, makes
it very easy to parallelize the downloading process and maximize the
download speed. Since the package management system only makes
available a hash of the entire content, the piece hashes will have
to be found through the P2P protocol. These hashes can be added to
the DHT, or requested from other peers.

@@

The DHT used is based on Khashmir, which is an implementation of the
Kademlia DHT \cite{kademlia} used by most of the existing BitTorrent
clients to implement trackerless operation. It has been modified to
better support the retrieval of multiple values per key, and to
improve the response time so as to satisfy users' demands for speed.
The values stored in the DHT are all bencoded dictionaries similar
to torrent files. Peers store their download location (IP address
and TCP port), as well as the piece strings (or hashes of piece
strings) as described in the previous section. Downloading is
accomplished by simple HTTP requests to the peers' identified to
have the desired file. All peers support HTTP/1.1, both in the
server and the client, which allows for pipelining of multiple
requests to a peer, and the requesting of smaller pieces of a large
file using the Range request header.

We have chosen a piece size of 512 kB, which means that 17,515
(79\%) of the almost 23 thousand packages will not require piece
information as they are smaller than a single piece. There are 3054
packages (13\%) that will require 2 to 4 pieces, for which the piece
hashes are stored directly with the peer download information in the
DHT. There are 1667 packages (7\%) that will require 5 to 70 pieces,
and so use a separate lookup in the DHT for the piece hashes.
Finally, there are only 62 packages (0.3\%) that require more than
70 pieces, and so for them a separate request to a peer for the
piece hashes is needed.

We have deployed the DHT from this implementation on PlanetLab to
test it's speed and effectiveness at delivering results. We find
that, even on the overloaded PlanetLab network, the responses to
lookups of keys take less than 10 seconds on average. Since each
package download requires a lookup before it can proceed, it may
take at least 10 seconds to process a request. However, due to
\texttt{apt} pipelining up to 10 requests at a time to our proxy,
the user will see an average response time (after a short startup)
of under a second, which should be responsive enough to satisfy most
users.

>>>>

\title{Leveraging Altruistic Peers to Reduce the Bandwidth Costs of Free Content Downloads}
\author{\IEEEauthorblockN{Cameron Dale}
\IEEEauthorblockA{School of Computing Science (student)\\
Simon Fraser University\\
Burnaby, British Columbia, Canada\\
Email: camerond@cs.sfu.ca}
\and
\IEEEauthorblockN{Jiangchuan Liu}
\IEEEauthorblockA{School of Computing Science (faculty)\\
Simon Fraser University\\
Burnaby, British Columbia, Canada\\
Email: jcliu@cs.sfu.ca}}

@@

@@

There are a large number of free content distributors using custom
package distribution systems over the Internet to distribute content
to their users: most Linux distributions (e.g. Debian, Red Hat,
Gentoo), Cygwin, CPAN, etc.
These distributors have developed many different
methods for this distribution, but they almost exclusively use a
client-server model to satisfy user requests. The contents' size,
popularity and number of users results in a large number of
requests, which usually requires a complicated network of mirrors to handle. Due
to the free nature of this content, many users are willing and able
to contribute upload bandwidth to this distribution, but currently
have no way to do this.

There are many peer-to-peer implementations available today, but
none is very well suited to this specific problem. Many distributors
make their content available in some form using BitTorrent \cite{COHEN03}, but it
is not ideally suited to an application such as this one. BitTorrent
is designed for large static collections of files, in which all
peers are interested in most or all of the content. But the packages
in this case are small, constantly being updated, and users are only
interested in a small subset of them. Other proposed implementations
suffer from some of the same, and other, problems and so none has
been well-used to address this problem.

We present a new peer-to-peer distribution model to meet these
demands. It is based on the lessons learned from many previous
implementations of successful peer-to-peer protocols, such as
Distributed Hash Tables and BitTorrent.
The model relies on the pre-existence of cryptographic
hashes of the content, which uniquely identify it for a request from
other peers. If the peer-to-peer download fails, then the original
request to the server is used as a fallback to prevent dissatisfied
users. The peer can then share this new content with others through
the P2P system. This system would efficiently satisfy the demands of
a large number of users, and should significantly reduce the
currently substantial bandwidth costs of hosting this content.

We also present a sample implementation based on the Debian package
distribution system. This implementation is simple, makes use of
mostly well-known technologies, and so serves as an example for
other free content distributors of the opportunity that can be easily met
with such a system.

@@

There are a large number of groups using the Internet to distribute
their free content. This content is made available from a free
download site, which usually requires a network of mirrors to
support the large number of requests. This content almost always
supports some kind of file verification using cryptographic hashes
to verify the download's accuracy or authenticity.

Most Linux distributions use a software package management system
that fetches packages to be installed from a network of mirrors.
Debian-based distributions uses the \texttt{apt} program, which
downloads Debian packages from one of many mirrors. RPM-based
distributions use \texttt{yum}, and Gentoo uses \texttt{portage},
which both operate in a similar way. Other free software
distributors also use a similar system: CPAN distributes files
containing software packages for the PERL programming language,
using SOAP RPC requests to find and download files; Cygwin provides
many of the standard Unix/Linux tools in a Windows environment,
using a package management tool that requests packages from
websites; two software distribution systems exist for Mac OSX,
fink and MacPorts, that also retrieve packages in this way. Also,
some systems use direct web downloads, but with a hash verification
file also available for download next to the desired file. These
hash files usually have the same file name, but with an added
extension identifying the hash used (e.g. \texttt{.md5} for the MD5
hash). Finally, there are other programs that make use of
cryptographic hashing to identify files: e.g. Git is a version
control system in which all files and changes are identified by
their SHA1 hash.

These systems all share some commonalities: the content is avaliable
for anyone to freely download, the content is divided into distinct
units (packages), users are typically not interested in downloading
all of the content available, and hashes of the content are
available before the download is attempted. We also expect that
there are many users of these systems that are motivated by
altruism to want to help out with this distribution. This is common
in these systems due to the free nature of the content being
delivered.

@@

This situation presents a clear opportunity to use a new
peer-to-peer file sharing protocol.
A downloading peer can lookup the package hash using the protocol and, if it
is found, download the file from those peers and verify the content.
If the package hash can not be found, the peer will
fallback to downloading from the original content location, and once
complete will announce to other peers indicating that it now has
the content. The original servers or mirrors thus act as \emph{seeds} for the
P2P system without any modification to them.

This functionality could be directly integrated into the
package management software, although this would be
difficult as the protocol should be running at all times, whereas the
package software typically only runs until the download request is complete.
Alternatively, since many of the package management software
implementations use HTTP requests to download the files, it
it possible to implement the P2P aspect as as HTTP caching
proxy. The proxy will get uncached requests first from the P2P system,
and then fallback to the normal HTTP request from a server. For
methods that don't use HTTP requests, other types of proxies
(possibly custom-made) may also be possible.

The sparse interest in a large
number of packages undergoing constant updating is well suited to
the functionality provided by a Distributed Hash Table (DHT). DHTs
require unique keys to store and retrieve strings of data, which the
cryptographic hashes used by the package management systems are
perfect for. The stored and retrieved strings can then be pointers
to the peers that have the content package that hashes to that key,
as well as any other information needed by the protocol to facilitate
the download.

To download larger files efficiently from a number of peers,
the file needs to be broken up into pieces, each with its own hash.
This method, inspired by BitTorrent, makes it very easy to
parallelize the downloading process and maximize the download speed.
Since the package management systems only make
available a hash of the entire content, the piece hashes will have
to be found through the P2P protocol.

@@

The DHT used is based on Khashmir, which is an implementation of
Kademlia \cite{kademlia} used by most of the existing BitTorrent
clients to implement trackerless operation. It has been modified to
better support the retrieval of multiple values (peers) per key, and to
improve the response time so as to satisfy users' demands for speed.
The values stored in the DHT are all bencoded dictionaries similar
to torrent files. Peers store their download location (IP address
and TCP port), as well as the piece strings (or hashes of piece
strings) as described in the previous section. Downloading is
accomplished by simple HTTP requests for the hash of the file,
which is sent to the peers identified from the DHT lookup to
have the desired file. All peers support HTTP/1.1, both as
servers and clients, which allows for pipelining of multiple
requests to a peer, and the requesting of smaller pieces of a large
file using the Range request header.

We have chosen a piece size of 512 kB, which means that 17,515
(79\%) of the almost 23,000 Debian packages will not require piece
information as they are smaller than a single piece. There are 3054
packages (13\%) that will require 2 to 4 pieces, for which the piece
hashes are stored directly with the peer download information in the
DHT. There are 1667 packages (7\%) that will require 5 to 70 pieces,
and so use a separate lookup in the DHT for the piece hashes.
Finally, there are only 62 packages (0.3\%) that require more than
70 pieces, and so need a separate request to a peer for the
piece hashes.

We have deployed the DHT part of this implementation on PlanetLab to
test it's speed and effectiveness at delivering results. We find
that, even on the overloaded PlanetLab network, the responses to
lookups of keys take less than 10 seconds on average. Since each
package download requires a lookup before it can proceed, it may
take at least 10 seconds to process a request. However, due to
\texttt{apt} pipelining up to 10 requests at a time to our proxy,
the user will see an average response time (after a short startup)
of under a second, which should be responsive enough to satisfy most
users.
